Clustering Beat-Chroma Patterns\\ in a Large Music Database

### 2010-07-26 Paper/Poster/poster.tex
getting there, some feedback from Ron included
final

\usepackage{epsfig}
\usepackage{multicol}
\usepackage{pstricks,pst-grad}
\usepackage{url}
\usepackage{wrapfig}
\usepackage{algorithm,algorithmic}

@@

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Can we infer anything about the underlying structure and common
features of commercial pop music by analyzing very large collections
of music audio?
\item Our interest: tonal content of the music -- i.e. the
harmony and melody.
\item Beat-synchronous chromagrams: rich enough
to generate musically-relevant results, simplified enough to
abstract away instrumentation and other stylistic details.
\end{list}

@@

%
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item \textbf{Goal}: identify meaningful information
about the musical structure represented in the entire database by
examining individual entries in this codebook.
\item \textbf{Method}: identify common patterns in beat-synchronous
chromagrams by learning codebooks from a large set of examples.
Individual codewords consist of short beat-chroma patches of
between 1 and 8 beats, optionally aligned to bar boundaries.
\item Prior work: ``shingles'' of \cite{Casey2007}, 
beat-synchronous analysis to identify the chorus by \cite{Bartsch2001},
and cover recognition by \cite{Ellis2007a}.
\end{list}

@@

\textbf{Chromagrams from the Echo Nest online API.}
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Feature analysis based on Echo Nest analyze API \cite{EchoNest}.
\item For any song, EchoNest provides a chroma
vector (length $12$) for every music event (called ``segment''), and a
segmentation of the song into beats and bars.
\item Averaging per-segment chroma over beat times results in a
beat-synchronous chroma feature representation.
\item Dataset size: $43,000$ songs.
\end{list}

\vspace{.5cm}
\textbf{Beat-Chroma Patches.}
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item We use Echo Nest analysis to break a song into a collection of 
beat-chroma ``patches'', typically one or two bars in length.
\item $82\%$ of the bars in our data were 4 beats long.
\item Other cases: we resample patches to a fixed length of 4 beats.
\item We rotate patches so that the first row contains the most energy.
This is analogous to key normalization, but
each patch is normalized independently.
\end{list}

@@

\myfig[0]{song_encoded.ps}{.7}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
Exemple of an encoded song (Good Day Sunshine
by \textit{The Beatles}) using a codebook. Codebook size:
$200$, codewords: $2$ bars.
\end{small}
\end{minipage}
\end{center}

\myfig[0]{codes_8bars.ps}{.7}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
Longer codewords spanning $8$ bars, randomly selected from
a $100$-entry codebook. Most patterns consist of sustained notes 
or chords, but code 0 shows one-bar alternations between 
two chords, and code 4 contains two cycles of a
$1\rightarrow1\rightarrow1\rightarrow2$ progression.

@@

Below are samples from actual codebooks and centroids.

\myfig[0]{close_patterns1.ps}{.8}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
Cluster around first centroid (in \textit{introduction} section).
Random selection from a cluster of size $639$. 
This gives an intuition of the
variance in one cluster.
\end{small}
\end{minipage}
\end{center}

\myfig[0]{codes1.ps}{.8}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
The 25 codes most commonly used. 
Codes are from a $200$-entry codebook trained
on $2$ bars, $12 × 8$ patches. The proportion of patches accounted
for by each pattern is shown in parentheses.
\end{small}
\end{minipage}
\end{center}

\myfig[0]{close_patterns_diff.ps}{.8}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
First three patterns from figure above ($2$nd row) presented
with the centroid ($1$st row) and the absolute difference between 
both ($3$ row).
\end{small}
\end{minipage}
\end{center}

@@

\mbox{} % dumb, but need something before vspace
\vspace{-1cm}
\myfig[0]{codes_lle.ps}{.8}

@@

\begin{wrapfigure}{r}{0.6\textwidth}
  \vspace{-20pt}
  \begin{center}
    \includegraphics[width=0.58\textwidth]{conf_mat_per_artist.ps}
  \end{center}
  \vspace{-20pt}
\end{wrapfigure}

\textbf{Artist recognition task}.
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item {\it artist20} data set: $1402$ songs from $20$ artists.
\item Baseline: GMM on MFCC ($59\%$), GMM on chromas ($33\%$),
random ($5\%$).
\item We get an accuracy of $\mathbf{23.4\%}$.
\end{list}
%We use the {\it artist20} data set: $1402$ songs from $20$ artists, 
%mostly rock and pop of different subgenres.  
%Previously published results using GMMs on MFCC features achieve an 
%accuracy of $59\%$, whereas using only chroma as a representation yields an
%accuracy of $33\%$  \cite{Ellis2007}.
%
%We get an accuracy of $\mathbf{23.4\%}$, random baseline 
%is around $5\%$. 
%The confusion matrix is shown here, note 
%that certain artists are recognized at an accuracy far above the average.

@@

% FOOTNOTE
\vspace{1cm}
\begin{minipage}{\textwidth}
\begin{small}
\begin{flushright}
Acknowledgements: thanks to G. Grindlay for
helpful comments. TBM is supported by a
NSERC scholarship. This material is based upon work
supported by IMLS grant LG-06-08-0073-08 and by NSF
grant IIS-0713334. \\
%Any opinions, findings and conclu-
%sions or recommendations expressed in this material are
%those of the authors and do not necessarily reflect the views
%of the sponsors.\\
Poster presented at ISMIR 2010 (Utrecht) for the paper:
Clustering Beat-Chroma Patterns in a Large Music Database by
T. Bertin-Mahieux, R. J. Weiss and D. P. W. Ellis. Contact:
\url{tb2332@columbia.edu}, code: \url{http://www.columbia.edu/~tb2332/}.
\end{flushright}
\end{small}
\end{minipage}

>>>>

\usepackage{epsfig}
\usepackage{multicol}
\usepackage{pstricks,pst-grad}
\usepackage{url}
\usepackage{wrapfig}
\usepackage{algorithm,algorithmic}
%\usepackage{times} % recommended by Ron

@@

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Availability of very large collections of music.
\item Can we infer anything about the underlying structure and common
features of pop music?
\item Our interest: tonal content of the music -- i.e. the
harmony and melody.
\item Beat-synchronous chromagrams: rich enough
to generate musically-relevant results,
abstract away instrumentation and stylistic details.
\end{list}

@@

%
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item \textbf{Goal}: identify meaningful information
about the musical structure
\item \textbf{Method}: identify common patterns in beat-synchronous
chromagrams by learning codebooks from a large set of examples.
\item Entire database
represented by individual entries in this codebook.
\item Individual codewords consist of short beat-chroma patches of
between 1 and 8 beats, optionally aligned to bar boundaries.
\item Prior work: ``shingles'' of \cite{Casey2007}, 
beat-synchronous analysis to identify the chorus by \cite{Bartsch2001},
and cover recognition by \cite{Ellis2007a}.
\end{list}

@@

\textbf{Chromagrams from the Echo Nest online API.}
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Feature analysis based on Echo Nest analyze API \cite{EchoNest}.
\item For any song, EchoNest provides a chroma
vector (length $12$) for every music event (called ``segment'')
\item Also provides a
segmentation of the song into beats and bars.
\item Averaging per-segment chroma over beat times results in a
beat-synchronous chroma feature representation.
\item Dataset size: $43,000$ songs.
\end{list}

\vspace{.5cm}
\textbf{Beat-Chroma Patches.}
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item We use Echo Nest analysis to break a song into a collection of 
beat-chroma ``patches'', typically one or two bars in length.
\item $82\%$ of the bars in our data were 4 beats long.
\item Other cases: we resample patches to a fixed length of 4 beats.
\item We rotate patches so that the first row contains the most energy
(this is analogous to key normalization).
\end{list}

@@

\myfig[0]{song_encoded.ps}{.68}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
Example of an \textbf{encoded song} (Good Day Sunshine
by \textit{The Beatles}) using a codebook. Codebook size:
$200$, codewords: $2$ bars.
\end{small}
\end{minipage}
\end{center}

\myfig[0]{codes_8bars.ps}{.68}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
Longer \textbf{codewords spanning $\mathbf{8}$ bars}, randomly selected from
a $100$-entry codebook. Most patterns consist of sustained notes 
or chords, but code 0 shows one-bar alternations between 
two chords, and code 4 contains two cycles of a
$1\rightarrow1\rightarrow1\rightarrow2$ progression.

@@

Below are samples from \textbf{actual codebooks and centroids}.

\myfig[0]{close_patterns1.ps}{.8}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
\textbf{Cluster around first centroid} (in \textit{introduction} section).
Random selection from a cluster of size $639$. 
This gives an intuition of the
variance in one cluster.
\end{small}
\end{minipage}
\end{center}

\myfig[0]{codes1.ps}{.8}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
The 25 \textbf{codes most commonly used}. 
Codes are from a $200$-entry codebook trained
on $2$ bars, $12 × 8$ patches. The proportion of patches accounted
for by each pattern is shown in parentheses.
\end{small}
\end{minipage}
\end{center}

\myfig[0]{close_patterns_diff.ps}{.8}
% figure caption
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{small}
First three patterns from figure above ($2$nd row) presented
with the centroid ($1$st row) and the \textbf{absolute difference} between 
patterns and centroid ($3$ row).
\end{small}
\end{minipage}
\end{center}

@@

\mbox{} % dumb, but need something before vspace
\vspace{-1cm}
\myfig[0]{codes_lle.ps}{.9}

@@

%\begin{wrapfigure}{r}{0.5\textwidth}
%  \vspace{-20pt}
%  \begin{center}
%    \hspace{-1cm}
%    \includegraphics[width=0.52\textwidth]{conf_mat_per_artist.ps}
%  \end{center}
%  \vspace{-20pt}
%\end{wrapfigure}

\textbf{Artist recognition task}.
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item {\it artist20} data set: $1402$ songs from $20$ artists.
\item Baseline: GMM - MFCC $59\%$, GMM - chromas $33\%$,
random $5\%$.
\item We get an accuracy of $\mathbf{23.4\%}$.
\end{list}

\myfig[0]{conf_mat_per_artist.ps}{.55}

%We use the {\it artist20} data set: $1402$ songs from $20$ artists, 
%mostly rock and pop of different subgenres.  
%Previously published results using GMMs on MFCC features achieve an 
%accuracy of $59\%$, whereas using only chroma as a representation yields an
%accuracy of $33\%$  \cite{Ellis2007}.
%
%We get an accuracy of $\mathbf{23.4\%}$, random baseline 
%is around $5\%$. 
%The confusion matrix is shown here, note 
%that certain artists are recognized at an accuracy far above the average.

@@

% FOOTNOTE
\vspace{1.5cm}
\begin{minipage}{\textwidth}
\begin{small}
\begin{flushright}
Acknowledgements: thanks to G. Grindlay for
helpful comments. TBM is supported by a
NSERC scholarship. This material is based upon work
supported by IMLS grant LG-06-08-0073-08 and by NSF
grant IIS-0713334. \\
%Any opinions, findings and conclu-
%sions or recommendations expressed in this material are
%those of the authors and do not necessarily reflect the views
%of the sponsors.\\
Poster presented at ISMIR 2010 (Utrecht) for the paper:
Clustering Beat-Chroma Patterns in a Large Music Database by
T. Bertin-Mahieux, R. J. Weiss and D. P. W. Ellis. Contact:
\url{tb2332@columbia.edu}, code: \url{http://www.columbia.edu/~tb2332/}.
\end{flushright}
\end{small}
\end{minipage}
